Articulation Analysis: The Sound of Teaching Music 1

### 2020-02-07 script/RNotebook/articulation.Rmd
correct plots, typos
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
update kot/kor analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 2'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for articulation.

@@

### Main predictions
Participants will exaggerate legato and staccato more while teaching.

@@

# SubNr as a factor
df_ioi$SubNr <- as.factor(df_ioi$SubNr)

@@

# Calculate KOR for each interval

df_kor <- data.frame()
for (subnr in unique(df_kot$SubNr)){
  for (block in unique(df_kot$BlockNr[df_kot$SubNr == subnr])){
    for (trial in unique(df_kot$TrialNr[df_kot$SubNr == subnr & df_kot$BlockNr == block])){
      df_current <- df_kot %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
      df_current$KOR <- df_current$KOT/ioi$Mean[ioi$SubNr == subnr & ioi$BlockNr == block & ioi$TrialNr == trial]
      df_kor <- rbind(df_kor, df_current)
    }
  }
}

@@

## Key-Overlap Time (KOT) - Articulation
### 1. All - individual{#kot_art_ind_all}
- Average KOT for each participant.

```{r kot_art, echo = FALSE}
# For each individual
kot_art <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_art <- cbind(kot_art[,1:4], as.data.frame(kot_art[,5]))
# Change colnames
colnames(kot_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kot_art)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

@@

### 2. All - group{#kot_art_group_all}
- Average KOT for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

@@

#### Plot
##### Bar chart
- Error bars show SEM of the mean

@@

##### Box plot

```{r kot_art_all_box, echo = FALSE}
p_kot_art_all_box <- ggplot(data = kot_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_all_box
```

#### Statistics
##### TBC

### 3. Trial by trial - individual{#kot_art_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_art_trial, echo = FALSE}
# For each individual
kot_art_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_art_trial <- cbind(kot_art_trial[,1:5], as.data.frame(kot_art_trial[,6]))
# Change colnames
colnames(kot_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kot_art_trial)
```

#### Plot
##### Line graph
```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 8}
p_kot_art_trial_legato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_legato
```

@@

### 4. Trial by trial - group{#kot_art_group_trial}
- Average KOT for each trial for each condition.

@@

## Key-Overlap Ratio (KOR) - Articulation
### 1. All - individual{#kor_art_ind_all}
- Average KOR for each participant.

```{r kor_art, echo = FALSE}
# For each individual
kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Subcomponent != "LtoS" & df_kot_art$Subcomponent != "StoL"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_art <- cbind(kor_art[,1:4], as.data.frame(kor_art[,5]))
# Change colnames
colnames(kor_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kor_art)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

@@

### 2. All - group{#kor_art_group_all}
- Average KOR for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

@@

##### Box plot

```{r kor_art_all_box, echo = FALSE}
p_kor_art_all_box <- ggplot(data = kor_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_all_box
```

#### Statistics
##### TBC

### 3. Trial by trial - individual{#kor_art_ind_trial}
- Average KOR for each trial for each participant.

@@

#### Plot
##### Line graph
```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 8}
p_kor_art_trial_legato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_legato
```

@@

### 4. Trial by trial - group{#kor_art_group_trial}
- Average KOR d for each trial for each condition.

@@

###########
###########

## Key-Overlap Time (KOT) - Dynamics
### 1. All - individual{#kot_dyn_ind_all}
- Average KOT for each participant.

```{r kot_dyn, echo = FALSE}
# For each individual
kot_dyn <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_dyn <- cbind(kot_dyn[,1:4], as.data.frame(kot_dyn[,5]))
# Change colnames
colnames(kot_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kot_dyn)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

@@

### 2. All - group{#kot_dyn_group_all}
- Average KOT for each condition.

@@

##### Box plot

@@

#### Statistics
##### TBC

### 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_dyn_trial, echo = FALSE}
# For each individual
kot_dyn_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_dyn_trial <- cbind(kot_dyn_trial[,1:5], as.data.frame(kot_dyn_trial[,6]))
# Change colnames
colnames(kot_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kot_dyn_trial)
```

#### Plot
##### Line graph
```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
p_kot_dyn_trial_forte <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_forte
```

@@

### 4. Trial by trial - group{#kot_dyn_group_trial}
- Average KOT for each trial for each condition.

@@

## Key-Overlap Ratio (KOR) - Dynamics
### 1. All - individual{#kor_dyn_ind_all}
- Average KOR for each participant.

```{r kor_dyn, echo = FALSE}
# For each individual
kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_dyn <- cbind(kor_dyn[,1:4], as.data.frame(kor_dyn[,5]))
# Change colnames
colnames(kor_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kor_dyn)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean

@@

### 2. All - group{#kor_dyn_group_all}
- Average KOR for each condition.

@@

##### Box plot

```{r kor_dyn_all_box, echo = FALSE}
p_kor_dyn_all_box <- ggplot(data = kor_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_all_box
```

#### Statistics
##### TBC

### 3. Trial by trial - individual{#kor_dyn_ind_trial}
- Average KOR for each trial for each participant.

```{r kor_dyn_trial, echo = FALSE}
# For each individual
kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Subcomponent != "FtoP" & df_kot_dyn$Subcomponent != "PtoF"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_dyn_trial <- cbind(kor_dyn_trial[,1:5], as.data.frame(kor_dyn_trial[,6]))
# Change colnames
colnames(kor_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kor_dyn_trial)
```

#### Plot
##### Line graph
```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
p_kor_dyn_trial_forte <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_forte
```

@@

### 4. Trial by trial - group{#kor_dyn_group_trial}
- Average KOR for each trial for each condition.

@@

>>>>

---
title: 'Articulation Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for articulation.

- Last checked: 07/02/20

# Outline - Articulation (KOT, KOR)
## [Experimental Design](#design)

## [Articulation](#kot-articulation)

### KOT
- Individual results
  + [All](#kot_art_ind_all)
  + [Trial by trial](#kot_art_ind_trial)

- Group results and stats (if available)
  + [All](#kot_art_group_all)
  + [Trial by trial](#kot_art_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kot_art_order_group_all)

- [Sequence plot](#sequence_art)

### KOR
- Individual results
  + [All](#kor_art_ind_all)
  + [Trial by trial](#kor_art_ind_trial)

- Group results and stats (if available)
  + [All](#kor_art_group_all)
  + [Trial by trial](#kor_art_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kor_art_order_group_all)

## [Dynamics](#kot-dynamics)

### KOT
- Individual results
  + [All](#kot_dyn_ind_all)
  + [Trial by trial](#kot_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kot_dyn_group_all)
  + [Trial by trial](#kot_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kot_dyn_order_group_all)

- [Sequence plot](#sequence_dyn)

### KOR
- Individual results
  + [All](#kor_dyn_ind_all)
  + [Trial by trial](#kor_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kor_dyn_group_all)
  + [Trial by trial](#kor_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kor_dyn_order_group_all)

@@

## Main predictions
Participants will exaggerate legato and staccato more while teaching.

@@

@@

# Calculate KOR for each interval
df_kor <- data.frame()
for (subnr in unique(df_kot$SubNr)){
  for (block in unique(df_kot$BlockNr[df_kot$SubNr == subnr])){
    for (trial in unique(df_kot$TrialNr[df_kot$SubNr == subnr & df_kot$BlockNr == block])){
      df_current <- df_kot %>% dplyr::filter(SubNr == subnr & BlockNr == block & TrialNr == trial)
      df_current$KOR <- df_current$KOT/ioi$Mean[ioi$SubNr == subnr & ioi$BlockNr == block & ioi$TrialNr == trial]
      df_kor <- rbind(df_kor, df_current)
    }
  }
}

@@

# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
## 1. All - individual{#kot_art_ind_all}
- Average KOT for each participant.

```{r kot_art, echo = FALSE}
# For each individual
kot_art <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_art <- cbind(kot_art[,1:4], as.data.frame(kot_art[,5]))
# Change colnames
colnames(kot_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kot_art)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

@@

## 2. All - group{#kot_art_group_all}
- Average KOT for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

@@

### Plot
#### Bar chart
- Error bars show SEM of the mean

@@

#### Box plot
```{r kot_art_all_box, echo = FALSE}
p_kot_art_all_box <- ggplot(data = kot_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_all_box
```

### Statistics
#### Normality check
```{r normality1, echo = FALSE}
# Reduce unused factors
kot_art$Subcomponent <- factor(kot_art$Subcomponent)
kot_art_norm <- by(kot_art$Mean, list(kot_art$Condition, kot_art$Subcomponent), shapiro.test)
print(kot_art_norm)
```

#### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
kot_art_aov <- ezANOVA(
  data = subset(df_kot_art, df_kot_art$Subcomponent == "Legato" | df_kot_art$Subcomponent == "Staccato")
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_art_aov$ANOVA$F <- round(kot_art_aov$ANOVA$F, 4)
kot_art_aov$ANOVA$p <- round(kot_art_aov$ANOVA$p, 4)
print(kot_art_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
print(summary(kot_art_aov_2))
```

#### Post-hoc (direct comparison)
```{r post-hoc1, echo = FALSE}
kot_ttest_legato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"], paired = TRUE)
kot_ttest_staccato <- t.test(kot_art$Mean[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"], kot_art$Mean[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"], paired = TRUE)
# Effect size
kot_ttest_legato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Legato"], kot_art$Condition[kot_art$Subcomponent == "Legato"], paired = TRUE)
kot_ttest_staccato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Staccato"], kot_art$Condition[kot_art$Subcomponent == "Staccato"], paired = TRUE)

print(kot_ttest_legato)
print(kot_ttest_staccato)
print(kot_ttest_legato_cohend)
print(kot_ttest_staccato_cohend)
```

#### Post-hoc (pairwise t test)
```{r post-hoc2, echo = FALSE}
# Create a group label for a post-hoc test
kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Legato"] <- "Performing-Legato" 
kot_art$Group[kot_art$Condition == "performing" & kot_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Legato"] <- "Teaching-Legato"
kot_art$Group[kot_art$Condition == "teaching" & kot_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
# pairwise comparison
kot_art_pairwise <- pairwise.t.test(kot_art$Mean, kot_art$Group, paired = TRUE, p.adjust.method = "BH")
kot_art_pairwise$p.value <- round(kot_art_pairwise$p.value, 4)
print(kot_art_pairwise)
```

## 3. Trial by trial - individual{#kot_art_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_art_trial, echo = FALSE}
# For each individual
kot_art_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_art, df_kot_art$Interval != 8 & df_kot_art$Interval != 16 & df_kot_art$Interval != 24 & df_kot_art$Interval != 41 & df_kot_art$Interval != 49 & df_kot_art$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_art_trial <- cbind(kot_art_trial[,1:5], as.data.frame(kot_art_trial[,6]))
# Change colnames
colnames(kot_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kot_art_trial)
```

### Plot
#### Line graph
```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 8}
p_kot_art_trial_legato <- ggplot(data = subset(kot_art_trial, kot_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_legato
```

@@

## 4. Trial by trial - group{#kot_art_group_trial}
- Average KOT for each trial for each condition.

@@

# Key-Overlap Ratio (KOR) - Articulation
## 1. All - individual{#kor_art_ind_all}
- Average KOR for each participant.

```{r kor_art, echo = FALSE}
# For each individual
kor_art <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_art, df_kor_art$Interval != 8 & df_kor_art$Interval != 16 & df_kor_art$Interval != 24 & df_kor_art$Interval != 41 & df_kor_art$Interval != 49 & df_kor_art$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_art <- cbind(kor_art[,1:4], as.data.frame(kor_art[,5]))
# Change colnames
colnames(kor_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kor_art)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

@@

## 2. All - group{#kor_art_group_all}
- Average KOR for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

@@

#### Box plot
```{r kor_art_all_box, echo = FALSE}
p_kor_art_all_box <- ggplot(data = kor_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_all_box
```

### Statistics
#### Normality check
```{r normality2, echo = FALSE}
# Reduce unused factors
kor_art$Subcomponent <- factor(kor_art$Subcomponent)
kor_art_norm <- by(kor_art$Mean, list(kor_art$Condition, kor_art$Subcomponent), shapiro.test)
print(kor_art_norm)
```

#### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
kor_art_aov <- ezANOVA(
  data = subset(df_kor_art, df_kor_art$Subcomponent == "Legato" | df_kor_art$Subcomponent == "Staccato")
  , dv = .(KOR)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_art_aov$ANOVA$F <- round(kor_art_aov$ANOVA$F, 4)
kor_art_aov$ANOVA$p <- round(kor_art_aov$ANOVA$p, 4)
print(kor_art_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
print(summary(kor_art_aov_2))
```

#### Post-hoc (direct comparison)
```{r post-hoc3, echo = FALSE}
kor_ttest_legato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"], paired = TRUE)
kor_ttest_staccato <- t.test(kor_art$Mean[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"], kor_art$Mean[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"], paired = TRUE)
# Effect size
kor_ttest_legato_cohend <- cohen.d(kor_art$Mean[kor_art$Subcomponent == "Legato"], kor_art$Condition[kor_art$Subcomponent == "Legato"], paired = TRUE)
kor_ttest_staccato_cohend <- cohen.d(kor_art$Mean[kor_art$Subcomponent == "Staccato"], kor_art$Condition[kor_art$Subcomponent == "Staccato"], paired = TRUE)

print(kor_ttest_legato)
print(kor_ttest_staccato)
print(kor_ttest_legato_cohend)
print(kor_ttest_staccato_cohend)
```

#### Post-hoc (pairwise t test)
```{r post-hoc4, echo = FALSE}
# Create a group label for a post-hoc test
kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Legato"] <- "Performing-Legato" 
kor_art$Group[kor_art$Condition == "performing" & kor_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Legato"] <- "Teaching-Legato"
kor_art$Group[kor_art$Condition == "teaching" & kor_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
# pairwise comparison
kor_art_pairwise <- pairwise.t.test(kor_art$Mean, kor_art$Group, paired = TRUE, p.adjust.method = "BH")
kor_art_pairwise$p.value <- round(kor_art_pairwise$p.value, 4)
print(kor_art_pairwise)
```

## 3. Trial by trial - individual{#kor_art_ind_trial}
- Average KOR for each trial for each participant.

@@

### Plot
#### Line graph
```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 8}
p_kor_art_trial_legato <- ggplot(data = subset(kor_art_trial, kor_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_legato
```

@@

## 4. Trial by trial - group{#kor_art_group_trial}
- Average KOR d for each trial for each condition.

@@

# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
## 1. All - individual{#kot_dyn_ind_all}
- Average KOT for each participant.

```{r kot_dyn, echo = FALSE}
# For each individual
kot_dyn <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_dyn <- cbind(kot_dyn[,1:4], as.data.frame(kot_dyn[,5]))
# Change colnames
colnames(kot_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kot_dyn)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

@@

## 2. All - group{#kot_dyn_group_all}
- Average KOT for each condition.

@@

#### Box plot

@@

### Statistics
#### Normality check
```{r normality3, echo = FALSE}
# Reduce unused factors
kot_dyn$Subcomponent <- factor(kot_dyn$Subcomponent)
kot_dyn_norm <- by(kot_dyn$Mean, list(kot_dyn$Condition, kot_dyn$Subcomponent), shapiro.test)
print(kot_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
kot_dyn_aov <- ezANOVA(
  data = subset(df_kot_dyn, df_kot_dyn$Subcomponent == "Forte" | df_kot_dyn$Subcomponent == "Piano")
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_dyn_aov$ANOVA$F <- round(kot_dyn_aov$ANOVA$F, 4)
kot_dyn_aov$ANOVA$p <- round(kot_dyn_aov$ANOVA$p, 4)
print(kot_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
print(summary(kot_dyn_aov_2))
```

## 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_dyn_trial, echo = FALSE}
# For each individual
kot_dyn_trial <- aggregate(KOT~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kot_dyn, df_kot_dyn$Interval != 8 & df_kot_dyn$Interval != 16 & df_kot_dyn$Interval != 24 & df_kot_dyn$Interval != 41 & df_kot_dyn$Interval != 49 & df_kot_dyn$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kot_dyn_trial <- cbind(kot_dyn_trial[,1:5], as.data.frame(kot_dyn_trial[,6]))
# Change colnames
colnames(kot_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kot_dyn_trial)
```

### Plot
#### Line graph
```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
p_kot_dyn_trial_forte <- ggplot(data = subset(kot_dyn_trial, kot_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_forte
```

@@

## 4. Trial by trial - group{#kot_dyn_group_trial}
- Average KOT for each trial for each condition.

@@

# Key-Overlap Ratio (KOR) - Dynamics
## 1. All - individual{#kor_dyn_ind_all}
- Average KOR for each participant.

```{r kor_dyn, echo = FALSE}
# For each individual
kor_dyn <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_dyn <- cbind(kor_dyn[,1:4], as.data.frame(kor_dyn[,5]))
# Change colnames
colnames(kor_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(kor_dyn)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

@@

## 2. All - group{#kor_dyn_group_all}
- Average KOR for each condition.

@@

#### Box plot
```{r kor_dyn_all_box, echo = FALSE}
p_kor_dyn_all_box <- ggplot(data = kor_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_all_box
```

### Statistics
#### Normality check
```{r normality4, echo = FALSE}
# Reduce unused factors
kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent)
kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test)
print(kor_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
kor_dyn_aov <- ezANOVA(
  data = subset(df_kor_dyn, df_kor_dyn$Subcomponent == "Forte" | df_kor_dyn$Subcomponent == "Piano")
  , dv = .(KOR)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4)
kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4)
print(kor_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
print(summary(kor_dyn_aov_2))
```

## 3. Trial by trial - individual{#kor_dyn_ind_trial}
- Average KOR for each trial for each participant.

```{r kor_dyn_trial, echo = FALSE}
# For each individual
kor_dyn_trial <- aggregate(KOR~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_kor_dyn, df_kor_dyn$Interval != 8 & df_kor_dyn$Interval != 16 & df_kor_dyn$Interval != 24 & df_kor_dyn$Interval != 41 & df_kor_dyn$Interval != 49 & df_kor_dyn$Interval != 57),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
kor_dyn_trial <- cbind(kor_dyn_trial[,1:5], as.data.frame(kor_dyn_trial[,6]))
# Change colnames
colnames(kor_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(kor_dyn_trial)
```

### Plot
#### Line graph
```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 8}
p_kor_dyn_trial_forte <- ggplot(data = subset(kor_dyn_trial, kor_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_forte
```

@@

## 4. Trial by trial - group{#kor_dyn_group_trial}
- Average KOR for each trial for each condition.

@@

# KOT-Order-Articulation
- Investigate whether there is any order effect of the conditions.
